---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
# Load all package functions to be available in the vignette
devtools::load_all()
```

# Introduction to the Optics Package

## 1. Introduction

The `Optics` package provides a standardized toolkit for evaluating the performance of machine learning models on optical survey data. This vignette demonstrates a complete workflow, from ingesting raw model output to generating final performance metrics and visualizations.

!Autonomous Underwater Vehicle

First, let's load the `Optics` package and other useful libraries like `dplyr`.


``` r
library(Optics)
library(dplyr)
library(ggplot2)
```

## 2. Creating Sample Data

For this example, we will create sample data representing raw model detections and a corresponding ground truth dataset. In a real workflow, this data would be ingested using `read_kwcoco()` or `read_viame_csv()`, which return `OpticsDetections` S4 objects.

For this vignette, we'll construct the data frames manually and then create the S4 objects. Note that the `OpticsDetections` class requires a specific set of columns.


``` r
# Sample model detections with scores
model_detections_df <- tibble(
  video_id = "vid01",
  image_id = "img01",
  annotation_id = 1:8,
  category_name = c("Gadus morhua", "Gadus morhua", "Melanogrammus aeglefinus", "Gadus morhua", "Pollachius virens", "Gadus morhua", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus"),
  score = c(0.95, 0.85, 0.80, 0.65, 0.50, 0.92, 0.75, 0.60),
  frame_index = c(10, 10, 15, 20, 20, 10, 15, 15),
  model_name = c(rep("Model A", 5), rep("Model B", 3)),
  bbox_x = 0, bbox_y = 0, bbox_width = 1, bbox_height = 1 # Dummy columns for validation
)

# Sample ground truth detections (no scores)
truth_detections_df <- tibble(
  video_id = "vid01",
  image_id = "img01",
  annotation_id = 9:12,
  category_name = c("Gadus morhua", "Gadus morhua", "Gadus morhua", "Urophycis tenuis"),
  frame_index = c(10, 10, 20, 30),
  score = 1.0, # Dummy score for validation
  bbox_x = 0, bbox_y = 0, bbox_width = 1, bbox_height = 1 # Dummy columns for validation
)
```

## 3. Working with `OpticsDetections` Objects

Now, let's convert these sample data frames into the formal `OpticsDetections` S4 objects that the package uses for analysis.

``` r
# Create OpticsDetections objects for each model and for the truth data
model_a_obj <- OpticsDetections(
  data = filter(model_detections_df, model_name == "Model A"),
  source_file = "manual", ingest_format = "manual"
)
model_b_obj <- OpticsDetections(
  data = filter(model_detections_df, model_name == "Model B"),
  source_file = "manual", ingest_format = "manual"
)
truth_obj <- OpticsDetections(
  data = truth_detections_df,
  source_file = "manual", ingest_format = "manual"
)
```

## 4. Summarizing Performance by Threshold

A key task is to evaluate how a model performs at different confidence thresholds. The `summarize_performance_by_threshold()` function automates this.

```r
# Define the thresholds we want to test
thresholds_to_test <- seq(0.5, 1.0, by = 0.1)

# Analyze Model A
perf_model_a <- summarize_performance_by_threshold(
  model_detections = model_a_obj,
  truth_detections = truth_obj,
  by = c("video_id", "category_name"),
  thresholds = thresholds_to_test
) %>% mutate(model_name = "Model A")

# Analyze Model B
perf_model_b <- summarize_performance_by_threshold(
  model_detections = model_b_obj,
  truth_detections = truth_obj,
  by = c("video_id", "category_name"),
  thresholds = thresholds_to_test
) %>% mutate(model_name = "Model B")

# Combine into a single data frame
performance_summary <- bind_rows(perf_model_a, perf_model_b)
```

## 5. Visualizing Performance

With the summary data, we can now create plots to compare the models. The plotting functions are now S4 generics.

### Performance by Threshold Plot

The `plot_performance_by_threshold()` function visualizes the trade-offs between precision, recall, and F1-score.


```{r plot-performance, fig.cap="Precision, Recall, and F1-Score for Model A and Model B across different confidence thresholds."}
plot_performance_by_threshold(
  performance_summary,
  model_col = "model_name",
  title = "Model Performance Comparison"
)
```

### Count Comparison Scatterplot

To see how well the model counts match the truth counts at a *specific* threshold (e.g., 0.8), we can generate a scatterplot. The `calculate_maxn` generic works on both `OpticsDetections` objects and standard `data.frame`s.


```{r plot-scatterplot, fig.cap="Model vs. Truth MaxN counts at a 0.8 confidence threshold."}
# We can still use standard dplyr pipes. The calculate_maxn S4 generic will
# dispatch the correct method for a data.frame.
model_counts_at_08 <- model_detections_df %>%
  filter(score >= 0.8) %>%
  calculate_maxn(group_cols = "model_name")

truth_counts <- calculate_maxn(truth_detections_df)

# Align each model's counts to the truth separately
aligned_a <- align_counts(
  model_counts = filter(model_counts_at_08, model_name == "Model A"),
  truth_counts = truth_counts,
  by = c("video_id", "category_name")
) %>% mutate(model_name = "Model A")

aligned_b <- align_counts(
  model_counts = filter(model_counts_at_08, model_name == "Model B"),
  truth_counts = truth_counts,
  by = c("video_id", "category_name")
) %>% mutate(model_name = "Model B")

aligned_counts <- bind_rows(aligned_a, aligned_b)

# Generate the plot
plot_counts_scatterplot(
  aligned_counts,
  model_col = "model_name",
  title = "MaxN Counts at 0.8 Confidence"
)
```

## 6. Advanced Usage: Confusion Matrix

A confusion matrix helps diagnose classification errors. The `Optics` package provides `align_detections()` to match individual model predictions to ground truth objects and `plot_confusion_matrix()` to visualize the results. This is different from `align_counts()`, which works on aggregated data like MaxN.

First, we align the raw detections from `Model A` against the ground truth at a specific confidence threshold.


```{r plot-confusion-matrix, fig.cap="Confusion Matrix for Model A at a 0.5 confidence threshold."}
# Align individual detections at a 0.5 confidence threshold
aligned_detections <- align_detections(
  model_detections = model_a_obj,
  truth_detections = truth_obj,
  threshold = 0.5
)

# Generate and plot the confusion matrix
plot_confusion_matrix(
  aligned_detections,
  title = "Confusion Matrix for Model A (Threshold = 0.5)"
)
```

This vignette provides a basic overview of a standard workflow. The `Optics` package contains many other S4 methods for more in-depth analysis, including generating ROC curves and analyzing the drivers of model error.