---
title: "Introduction to the Optics Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the Optics Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 4.5
)
```

<!-- Note: This vignette uses functions that may require packages like 'lme4'. -->
<!-- Ensure such packages are listed under "Suggests:" in the DESCRIPTION file. -->

## Introduction

The `Optics` package provides a standardized toolkit for evaluating the performance of machine learning models on optical survey data. This vignette demonstrates a complete workflow, from ingesting raw model output to generating final performance metrics and visualizations.

First, let's load the `Optics` package and other useful libraries like `dplyr`.

```{r setup}
library(Optics)
library(dplyr)
library(ggplot2)
```

## 1. Creating Sample Data

For this example, we will create sample data representing raw model detections and a corresponding ground truth dataset.

*   **Model Detections**: Detections from two different models (`Model A` and `Model B`) across a range of confidence scores.
*   **Truth Detections**: The human-annotated ground truth data for a single video.

```{r create-data}
# Sample model detections with scores
model_detections <- tibble(
  video_id = "vid01",
  category_name = c("Gadus morhua", "Gadus morhua", "Melanogrammus aeglefinus", "Gadus morhua", "Pollachius virens"),
  score = c(0.95, 0.85, 0.80, 0.65, 0.50),
  frame_index = c(10, 10, 15, 20, 20),
  model_name = "Model A"
) %>%
  bind_rows(tibble(
    video_id = "vid01",
    category_name = c("Gadus morhua", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus"),
    score = c(0.92, 0.75, 0.60),
    frame_index = c(10, 15, 15),
    model_name = "Model B"
  ))

# Sample ground truth detections (no scores)
truth_detections <- tibble(
  video_id = "vid01",
  category_name = c("Gadus morhua", "Gadus morhua", "Gadus morhua", "Urophycis tenuis"),
  frame_index = c(10, 10, 20, 30)
)

knitr::kable(model_detections, caption = "Sample Model Detections")
knitr::kable(truth_detections, caption = "Sample Ground Truth Detections")
```

## 2. Summarizing Performance by Threshold

A key task is to evaluate how a model performs at different confidence thresholds. The `summarize_performance_by_threshold()` function automates this by calculating metrics like precision, recall, and F1-score for each threshold.

We will run this for each model separately and then combine the results.

```{r summarize-performance}
# Define the thresholds we want to test
thresholds_to_test <- seq(0.5, 0.9, by = 0.1)

# Analyze Model A
perf_model_a <- summarize_performance_by_threshold(
  model_detections = filter(model_detections, model_name == "Model A"),
  truth_detections = truth_detections,
  by = c("video_id", "category_name"),
  thresholds = thresholds_to_test
) %>% mutate(model_name = "Model A")

# Analyze Model B
perf_model_b <- summarize_performance_by_threshold(
  model_detections = filter(model_detections, model_name == "Model B"),
  truth_detections = truth_detections,
  by = c("video_id", "category_name"),
  thresholds = thresholds_to_test
) %>% mutate(model_name = "Model B")

# Combine into a single data frame
performance_summary <- bind_rows(perf_model_a, perf_model_b)

knitr::kable(performance_summary, caption = "Performance Metrics by Model and Threshold", digits = 2)
```

## 3. Visualizing Performance

With the summary data, we can now create plots to compare the models.

### Performance by Threshold Plot

The `plot_performance_by_threshold()` function is perfect for visualizing the trade-offs between precision, recall, and F1-score. This helps in selecting an optimal confidence threshold for each model.

```{r plot-performance, fig.cap="Precision, Recall, and F1-Score for Model A and Model B across different confidence thresholds.", out.width="100%"}
plot_performance_by_threshold(
  performance_summary,
  model_col = model_name,
  title = "Model Performance Comparison"
)
```

### Count Comparison Scatterplot

To see how well the model counts match the truth counts at a *specific* threshold (e.g., 0.8), we can generate a scatterplot.

```{r plot-scatterplot, fig.cap="Model vs. Truth MaxN counts at a 0.8 confidence threshold.", out.width="100%"}
# First, calculate counts at a chosen threshold
model_counts_at_08 <- model_detections %>%
  filter(score >= 0.8) %>%
  calculate_maxn(group_cols = "model_name")

truth_counts <- calculate_maxn(truth_detections)

# Align the data
aligned_counts <- align_counts(
  model_counts = model_counts_at_08,
  truth_counts = truth_counts,
  by = c("video_id", "category_name", "model_name")
)

# Generate the plot, faceting by model
plot_counts_scatterplot(
  aligned_counts,
  model_col = model_name,
  title = "MaxN Counts at 0.8 Confidence"
)
```

This vignette provides a basic overview of a standard workflow. The `Optics` package contains many other functions for more in-depth analysis, including generating ROC curves (`plot_roc_curve`), multi-class confusion matrices (`plot_multiclass_confusion_matrix`), and analyzing the drivers of model error (`analyze_performance_drivers`).