---
title: "VIAME Output Analysis Report"
author: "Author: Kelsey Martin \n\n VIAME Analysts: Jack Prior and Sara Thomas"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: readable  # A clean theme
    toc: true        # Adds a table of contents
    toc_float: true  # Makes the TOC float on the side
    toc_depth: 2     # Tell the TOC which header level to display
    fig_caption: yes
params:
  species: !r NULL
  outdir: !r NULL
  remove_large_schools: "n"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6)
library(Optics)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(ggpubr) # For stat_regline_equation
library(stringr)
library(broom)
library(flextable)
library(MuMIn)
library(lsmeans)
library(boot)
library(pander)
library(purrr)
options(width = 850)
opts_knit$set(progress = FALSE, verbose = FALSE)
panderOptions('table.split.table', Inf)
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.duplicate.label = "allow")

did_we_skip <- "no"

get_script_dir <- function() {
  cmd_args <- commandArgs(trailingOnly = FALSE)
  file_arg <- grep("^--file=", cmd_args, value = TRUE)
  if (length(file_arg) > 0) {
    return(dirname(normalizePath(sub("^--file=", "", file_arg[1]))))
  }
  if (!is.null(sys.frames()[[1]]$ofile)) {
    return(dirname(normalizePath(sys.frames()[[1]]$ofile)))
  }
  normalizePath(getwd())
}

script.dir <- get_script_dir()
data_root_candidates <- c(
  file.path(script.dir, "Data"),
  file.path(script.dir, "..", "..", "..", "Data"),
  file.path(script.dir, "..", "..", "Data"),
  file.path(getwd(), "Data")
)
data_root <- data_root_candidates[dir.exists(data_root_candidates)][1]
if (is.na(data_root)) {
  stop("Could not find a Data directory. Expected one of: ", paste(data_root_candidates, collapse = ", "))
}
wrkdir <- data_root
track_dir_candidates <- c(
  file.path(data_root, "Tracks"),
  file.path(data_root, "GFISHER"),
  file.path(data_root, "SEFSC")
)
trkdir <- track_dir_candidates[dir.exists(track_dir_candidates)][1]
trthdir <- file.path(data_root, "Truth")
outdir <- if (!is.null(params$outdir) && nzchar(params$outdir)) {
  params$outdir
} else {
  file.path(script.dir, "Output")
}

dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "Part I - Counts"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "Part II - Groundtruthing"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "Part III - Data Analysis"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "Part III - Data Analysis", "Figures"), recursive = TRUE, showWarnings = FALSE)

Allreadtimes <- read.csv(file.path(wrkdir, "All GFISHER Read Times.csv"))
Readtimekey <- read.csv(file.path(wrkdir, "final_filled_key_v2.csv")) %>%
  mutate(Videotime = sub("\..*", "", Timestamp))
Species_List <- read.csv(file.path(wrkdir, "Species List.csv"))
fwri_ref_key <- read.csv(file.path(wrkdir, "env3LABS_93to24.csv")) %>%
  dplyr::mutate(
    SITE_ID = gsub("_|-", "", SITE_ID),
    REFERENCE = gsub("_|-", "", REFERENCE),
    Deployment = ifelse(YEAR < 2024 | LAB == "FWRI", REFERENCE, SITE_ID)
  )

track_files <- if (dir.exists(trkdir)) {
  list.files(trkdir, pattern = "_tracks.*\.csv$", full.names = TRUE, recursive = TRUE)
} else {
  character(0)
}

truth_candidates <- c(
  file.path(trthdir, "maxn3LABS_93to24.csv"),
  file.path(wrkdir, "maxn3LABS_93to24.csv"),
  file.path(trkdir, "maxn3LABS_93to24.csv")
)
truth_path <- truth_candidates[file.exists(truth_candidates)][1]

if (length(track_files) == 0) {
  stop("No SEFSC track files found in the expected GFISHER track directories.")
}
if (is.na(truth_path) || !file.exists(truth_path)) {
  stop("Missing truth file: maxn3LABS_93to24.csv in Data/Truth, Data/, or the track directory.")
}

extract_deployment_id <- function(path) {
  sub("_tracks.*$", "", basename(path))
}

normalize_reference_id <- function(x) {
  x %>%
    toupper() %>%
    str_replace("^(\d{4})-(N(?:CD|CO)-\d{3})$", "\1_\2")
}

normalize_compact_id <- function(x) {
  x %>%
    toupper() %>%
    gsub("_|-", "", .)
}

kwcoco_preview <- convert_track_csv_to_kwcoco(
  track_files[[1]],
  video_name = extract_deployment_id(track_files[[1]])
)

Allreadtimes$ReferenceID <- gsub("_|-", "", Allreadtimes$ReferenceID)
start_lookup <- Allreadtimes %>%
  left_join(Readtimekey, by = c("StartTime" = "Videotime")) %>%
  dplyr::group_by(ReferenceID) %>%
  dplyr::summarise(Start = min(Frame, na.rm = TRUE), .groups = "drop")
end_lookup <- Allreadtimes %>%
  left_join(Readtimekey, by = c("EndTime" = "Videotime")) %>%
  dplyr::group_by(ReferenceID) %>%
  dplyr::summarise(End = max(Frame, na.rm = TRUE), .groups = "drop")
frame_lookup <- full_join(start_lookup, end_lookup, by = "ReferenceID") %>%
  mutate(
    Start = ifelse(is.infinite(Start), NA, Start),
    End = ifelse(is.infinite(End), NA, End)
  )

stitched_model_df <- map2_dfr(track_files, extract_deployment_id(track_files), function(track_file, deployment_id) {
  detections <- read_viame_csv(track_file, video_id = deployment_id)

  detections@data %>%
    mutate(
      deployment_id = deployment_id,
      deployment_reference_id = normalize_reference_id(deployment_id),
      deployment_reference_compact = normalize_compact_id(deployment_reference_id),
      source_track_file = basename(track_file),
      year = as.numeric(str_extract(deployment_id, "^\d{4}"))
    )
})

stitched_model_df <- stitched_model_df %>%
  left_join(frame_lookup, by = c("deployment_reference_compact" = "ReferenceID")) %>%
  dplyr::filter((frame_index >= Start & frame_index <= End) | (is.na(Start) & is.na(End))) %>%
  dplyr::select(-any_of(c("Start", "End")))

stitched_model <- OpticsDetections(
  stitched_model_df,
  source_file = paste(basename(track_files), collapse = ", "),
  ingest_format = "viame_csv_stitched"
)

truth_maxn <- read_wide_maxn(truth_path, video_id_col = REFERENCE) %>%
  mutate(
    deployment_id = str_extract(video_id, "\d{4}[_-]N(?:CD|CO)-\d{3}"),
    deployment_id = ifelse(is.na(deployment_id), video_id, deployment_id),
    deployment_reference_id = normalize_reference_id(deployment_id),
    deployment_reference_compact = normalize_compact_id(deployment_reference_id),
    class_label = category_name,
    true_count = truth_count,
    year = as.numeric(str_extract(deployment_id, "^\d{4}"))
  )

fwri_lookup <- fwri_ref_key %>%
  transmute(
    reference_compact = normalize_compact_id(REFERENCE),
    site_compact = normalize_compact_id(SITE_ID)
  )

model_deployments <- unique(stitched_model@data$deployment_reference_compact)
model_lookup <- fwri_lookup %>%
  filter(site_compact %in% model_deployments | reference_compact %in% model_deployments)
truth_refs_to_keep <- unique(c(model_deployments, model_lookup$reference_compact, model_lookup$site_compact))

truth_aligned <- truth_maxn %>%
  filter(deployment_reference_compact %in% truth_refs_to_keep) %>%
  mutate(class_label = trimws(class_label))

deployment_label_lookup <- truth_aligned %>%
  distinct(deployment_reference_compact, deployment_reference_id)

model_maxn <- calculate_maxn(stitched_model) %>%
  transmute(
    deployment_reference_compact = normalize_compact_id(normalize_reference_id(video_id)),
    class_label_raw = trimws(category_name),
    maxn = maxn
  ) %>%
  left_join(
    Species_List %>%
      transmute(class_label_raw = trimws(Spec_Viame_Dash), class_label = trimws(Species)),
    by = "class_label_raw"
  ) %>%
  mutate(class_label = ifelse(is.na(class_label), class_label_raw, class_label))

model_vs_truth_aligned <- align_counts(
  model_counts = model_maxn %>%
    transmute(video_id = deployment_reference_compact, category_name = class_label, maxn = maxn),
  truth_counts = truth_aligned %>%
    transmute(video_id = deployment_reference_compact, category_name = class_label, true_count = true_count),
  by = c("video_id", "category_name"),
  model_col = maxn,
  truth_col = true_count
)

aligned_maxn <- model_vs_truth_aligned %>%
  rename(deployment_reference_compact = video_id) %>%
  left_join(deployment_label_lookup, by = "deployment_reference_compact") %>%
  mutate(
    deployment_reference_id = ifelse(
      is.na(deployment_reference_id),
      deployment_reference_compact,
      deployment_reference_id
    )
  )

model_vs_manual_maxn <- aligned_maxn %>%
  transmute(
    deployment_reference_id = deployment_reference_id,
    class_label = category_name,
    VIAME_MaxN = model_count,
    Manual_MaxN = truth_count,
    Difference = Manual_MaxN - VIAME_MaxN
  ) %>%
  arrange(deployment_reference_id, class_label)

total_comparisons <- nrow(distinct(aligned_maxn, deployment_reference_compact, category_name))
maxn_binary_metrics <- calculate_binary_metrics(
  aligned_maxn %>% rename(video_id = deployment_reference_compact),
  total_comparisons = total_comparisons
)

comparison_thresholds <- c(seq(0.1, 0.9, by = 0.1), 0.95)
aligned_threshold_runs <- purrr::map_dfr(comparison_thresholds, function(confidence_threshold) {
  threshold_model_df <- stitched_model@data %>%
    filter(score >= confidence_threshold)

  threshold_maxn <- if (nrow(threshold_model_df) == 0) {
    tibble(
      deployment_reference_compact = character(),
      class_label = character(),
      maxn = numeric(),
      year = numeric()
    )
  } else {
    calculate_maxn(threshold_model_df) %>%
      transmute(
        deployment_reference_compact = normalize_compact_id(normalize_reference_id(video_id)),
        class_label_raw = trimws(category_name),
        maxn = maxn,
        year = as.numeric(str_extract(video_id, "^\d{4}"))
      ) %>%
      left_join(
        Species_List %>%
          transmute(class_label_raw = trimws(Spec_Viame_Dash), class_label = trimws(Species)),
        by = "class_label_raw"
      ) %>%
      mutate(class_label = ifelse(is.na(class_label), class_label_raw, class_label)) %>%
      select(deployment_reference_compact, class_label, maxn, year)
  }

  aligned_threshold <- align_counts(
    model_counts = threshold_maxn %>%
      transmute(
        video_id = deployment_reference_compact,
        category_name = class_label,
        maxn = maxn,
        year = year
      ),
    truth_counts = truth_aligned %>%
      transmute(
        video_id = deployment_reference_compact,
        category_name = class_label,
        true_count = true_count,
        year = year
      ),
    by = c("video_id", "category_name"),
    model_col = maxn,
    truth_col = true_count
  )

  aligned_threshold %>%
    rename(deployment_reference_compact = video_id) %>%
    left_join(deployment_label_lookup, by = "deployment_reference_compact") %>%
    mutate(
      deployment_reference_id = ifelse(
        is.na(deployment_reference_id),
        deployment_reference_compact,
        deployment_reference_id
      ),
      year = dplyr::coalesce(year.x, year.y, as.numeric(str_extract(deployment_reference_id, "^\d{4}"))),
      Version = "Optics package pipeline",
      Confidence = confidence_threshold,
      Species = trimws(category_name),
      Deployment = deployment_reference_id,
      VIAME_MaxN = model_count,
      Manual = truth_count,
      Difference = Manual - VIAME_MaxN,
      Model_info = paste(year, Version, Confidence, sep = "_")
    ) %>%
    select(Deployment, year, Version, Confidence, Species, VIAME_MaxN, Manual, Difference, Model_info)
})

combined_master <- aligned_threshold_runs %>%
  distinct(Deployment, Species, Confidence, .keep_all = TRUE) %>%
  filter(!is.na(year))

if (nrow(combined_master) == 0) {
  stop("No deployment-level MaxN comparisons were generated from the stitched model tracks and REFERENCE data.")
}

calculate_metrics <- function(df, species = "none") {
  summarize_metrics <- function(.data) {
    .data %>% dplyr::summarize(
      Agree = ifelse(Manual == VIAME_MaxN, 1, 0),
      Difference = Manual - VIAME_MaxN,
      Relaxed = ifelse(Difference == 1 | Difference == -1 | Difference == 0, 1, 0),
      TP = sum(Manual > 0 & VIAME_MaxN > 0, na.rm = TRUE),
      FP = sum(Manual == 0 & VIAME_MaxN > 0, na.rm = TRUE),
      FN = sum(Manual > 0 & VIAME_MaxN == 0, na.rm = TRUE),
      TN = sum(Manual == 0 & VIAME_MaxN == 0, na.rm = TRUE),
      Precision = TP / (TP + FP),
      Recall_TPR = TP / (TP + FN),
      FPR = FP / (FP + TN),
      FNR = FN / (TP + FN),
      Accuracy = (TP + TN) / (TP + FP + FN + TN),
      False_P_Ratio = FP / (TP + FN + TN),
      False_N_Ratio = FN / (TP + FP + TN),
      Total_Actual_Positives = TP + FN,
      Total_Actual_Negatives = FP + TN,
      .groups = 'drop'
    )
  }

  if (species == "none") {
    df_out <- df %>%
      dplyr::group_by(year, Version, Confidence) %>%
      summarize_metrics()
  } else if (species == "all") {
    df_out <- df %>%
      dplyr::group_by(year, Version, Confidence, Species) %>%
      summarize_metrics()
  } else if (any(species %in% df$Species) == TRUE) {
    df_out <- df %>%
      dplyr::filter(Species == species) %>%
      dplyr::group_by(year, Version, Confidence, Species) %>%
      summarize_metrics()
  } else {
    print("No species detected with that name. Check spelling and try again.")
    return(NULL)
  }

  df_out %>%
    dplyr::mutate(across(where(is.numeric), ~ ifelse(is.nan(.), NA, .))) %>%
    dplyr::mutate(across(where(is.numeric), ~ ifelse(is.infinite(.), NA, .)))
}

percent_metric <- function(df, variable1, variable2, group, metric) {
  df %>%
    group_by({{ variable1 }}, {{ variable2 }}, {{ group }}) %>%
    summarise(
      percentage_1s = mean({{ metric }}) * 100,
      percentage_0s = (1 - mean({{ metric }})) * 100,
      count = n(),
      .groups = "drop"
    )
}

metrics <- calculate_metrics(combined_master, species = "all")
percent_agreement <- percent_metric(metrics, year, Species, Confidence, Agree)
relaxed_agreement <- percent_metric(metrics, year, Species, Confidence, Relaxed)
percent_agreement$Agreement <- "Exact"
relaxed_agreement$Agreement <- "Relaxed"
PA.comp <- bind_rows(percent_agreement, relaxed_agreement)

species_values <- sort(unique(combined_master$Species))
if (!is.null(params$species) && nzchar(params$species) && !params$species %in% species_values) {
  stop("Requested species not found in the stitched deployment-level MaxN comparison: ", params$species)
}
species <- if (!is.null(params$species) && nzchar(params$species)) params$species else species_values[[1]]
spec_pretty <- gsub("_", " ", str_to_sentence(species))
spec_master <- combined_master %>% dplyr::filter(Species == species)
sub_false <- metrics %>% dplyr::filter(Species == species)
pa_spec <- PA.comp %>% dplyr::filter(Species == species)
remove_large_schools <- if (!is.null(params$remove_large_schools) && nzchar(params$remove_large_schools)) params$remove_large_schools else "n"
unique_years <- sort(unique(spec_master$year))

count_models <- function(df, species, year, confidence) {
  df <- df %>%
    dplyr::filter(Species == species, year == year, Confidence == confidence)
  glm.out <- glm(Manual ~ VIAME_MaxN, family = poisson(), data = df)
  anova(glm.out)
  print(summary(glm.out))

  lm.out <- lm(Manual ~ VIAME_MaxN, data = df)
  anova(lm.out)
  print(summary(lm.out))
}

lm_plots <- function(df, species, year) {
  df <- df %>%
    dplyr::filter(Species == species, year == year)
  spec_title <- paste(gsub("_", " ", str_to_sentence(species)), "Linear Model")
  lm_plot <- ggplot(data = df, aes(x = Manual, y = VIAME_MaxN, group = Confidence))+
    stat_regline_equation(label.x = -Inf, label.y = Inf, vjust = 2, hjust = -0.05) +
    stat_regline_equation(label.x = -Inf, label.y = Inf, vjust = 3, hjust = -0.05, aes(label = after_stat(rr.label))) +
    scale_y_continuous(expand = c(0,0.1)) +
    geom_smooth(se = TRUE, method = lm) +
    geom_point(data = df, aes(Manual, VIAME_MaxN)) +
    geom_count(aes(size = after_stat(n), group = 1), pch = 21, fill = "grey40", color = "black") +
    scale_size_area(max_size = 6)  +
    facet_wrap(~Confidence, scales = "free_y") +
    xlab("Reference MaxN") + ylab("Model MaxN") + ggtitle(spec_title, subtitle = paste("Year:", year)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
                      plot.title = element_text(vjust = 4, size = 14, hjust = 0.5),
                      plot.subtitle = element_text(vjust = 4, size = 12, hjust = 0.5, face = "italic"),
                      plot.margin = unit(c(0.3,0.2,0.2,0.2), "inches"),
                      legend.title = element_text(size = 12, color = "grey20"),
                      text = element_text(size = 12, color = "grey20")) + guides(size = guide_legend(title = "Number of
Agreements"))

  ggsave(lm_plot, filename = paste0(outdir, "Part III - Data Analysis/Figures/", species, "_", year, "_allconf_lmplots.png"), device = "png", dpi = 300, height = 9, width = 11)
  lm_plot
}
```

```{css, echo=FALSE}
/* --- Change Header Colors --- */
h1 {
  color: #0085CA; 
  font-size: 30px;
}

h2 {
  color: #0072BB; 
}

h3 {
  color: #003087;
}

h4 {
  color: #002467; 
}

h5 {
  color: #0072BB; 
  font-size: 20px;
}
/* --- Change Tab Colors --- */
/* This changes the text color of INACTIVE tabs */
.nav-tabs > li > a {
  color: #5A5F63; /* Dark Grey */
}

/* This changes the text color of the ACTIVE tab */
.nav-tabs > li.active > a,
.nav-tabs > li.active > a:hover,
.nav-tabs > li.active > a:focus {
  color: #de425b !important; /* A bright red. !important overrides defaults */
  font-weight: bold;
}

/* --- Change Pill Colors (.tabset-pills) --- */
/* This changes the text color of INACTIVE pills */
.nav-pills > li > a {
  color: #5A5F63; /* Dark Grey */
}

/* This changes the BACKGROUND of the ACTIVE pill */
.nav-pills > li.active > a,
.nav-pills > li.active > a:hover,
.nav-pills > li.active > a:focus {
  color: #FFFFFF !important;            /* White text */
  background-color: #027779 !important; /* Your red as the background */
  font-weight: bold;
}
```
<br> 

# `r spec_pretty` - Introduction 

The report below details the deployment-level MaxN performance metrics for `r spec_pretty` from the Optics package workflow `r as.character(unique(spec_master$Version))` for the G-FISHER Pascagoula and Panama City labs. Video-level track files are stitched into deployment-level `OpticsDetections` objects before they are aligned against the manual `REFERENCE` counts, and frame windows are shortened to reflect the times during which deployments were read. Data preparation removes deployment/species combinations where both model and reference MaxN are zero while preserving the legacy per-species analysis workflow and figures. For this species, the deployment years span `r as.numeric(min(spec_master$year))` to `r as.numeric(max(spec_master$year))`. `r if(remove_large_schools == "y" | remove_large_schools == "Y") { "Large-school reference counts (i.e., 299, 399, or 999) have been removed from the summarized deployment-level comparison."}`

***

# Optimal Confidence Across Years
```{r, echo=FALSE, warning=FALSE, error=FALSE, include = FALSE, fig.height = 6, fig.width = 7}
metrics_group <- sub_false %>%
  dplyr::group_by(Confidence) %>%
  dplyr::summarise(FN = mean(FN), FP = mean(FP), Precision = mean(Precision), Accuracy = mean(Accuracy), .groups = "drop")

pa_group <- pa_spec %>%
  dplyr::group_by(Confidence, Agreement) %>%
  dplyr::summarise(`Percent Agreement` = mean(percentage_1s), .groups = "drop")

# If requested, filter out large-school reference counts before summarizing threshold performance.
if (remove_large_schools == "y" | remove_large_schools == "Y") {
  spec_master <- spec_master %>%
    dplyr::filter(!Manual %in% c(299, 399, 999))
} else {
  spec_master <- spec_master %>%
    dplyr::mutate(
      Manual = ifelse(Manual == 399 | Manual == 999, 299, Manual),
      VIAME_MaxN = ifelse(VIAME_MaxN > 75, 299, VIAME_MaxN)
    )
}

if ((nrow(sub_false) > 10 | sum(pa_group$`Percent Agreement`, na.rm = TRUE) != 0)) {
  pa_group_long <- pa_group %>%
    tidyr::pivot_wider(names_from = Agreement, values_from = `Percent Agreement`)

  params_group <- metrics_group %>%
    left_join(pa_group_long, by = "Confidence") %>%
    rename(
      `Avg False Negatives` = FN,
      `Avg False Positives` = FP,
      `Avg % Agreement (Exact)` = Exact,
      `Avg % Agreement (Relaxed)` = Relaxed
    )

  sum_precis <- list(c(1:9,95))
  for (c in c(1:9,95)) {
    con = as.numeric(paste0("0.", c))
    df <- spec_master %>% dplyr::filter(Confidence == con)
    if (nrow(df) > 1) {
      ap.out <- agePrecision(~Manual+VIAME_MaxN, data = df)
      sum_precis[[c]] <- data.frame(summary(ap.out, what = "precision")) %>%
        dplyr::mutate(Confidence = con) %>%
        dplyr::select(Confidence, ACV, PercAgree)
    } else {
      sum_precis[[c]] <- df %>%
        dplyr::mutate(Confidence = Confidence, ACV = NA, PercAgree = NA) %>%
        dplyr::select(Confidence, ACV, PercAgree)
    }
  }
  precis <- do.call(rbind, sum_precis)
  params_group <- params_group %>%
    left_join(precis, by = "Confidence") %>%
    dplyr::select(Confidence, `Avg False Negatives`, `Avg False Positives`, `Avg % Agreement (Exact)`, `Avg % Agreement (Relaxed)`, ACV, PercAgree) %>%
    dplyr::arrange(ACV, .desc = TRUE)
  params_plot_dat <- params_group %>%
    tidyr::pivot_longer(-Confidence, names_to = "Metric", values_to = "value")

  optim_goals <- tibble::tribble(
    ~Metric,                      ~Goal,      ~Weight,
    "Avg False Negatives",         "min",       0.1,
    "Avg False Positives",         "min",       0.25,
    "Avg % Agreement (Exact)",     "max",        0,
    "Avg % Agreement (Relaxed)",   "max",        0,
    "ACV",                         "min",       0.35,
    "PercAgree",                   "max",       0.3
  )

  best_confidence_values <- params_plot_dat %>%
    inner_join(optim_goals, by = "Metric") %>%
    group_by(Metric, Goal) %>%
    dplyr::filter(
      (Goal == "min" & value == min(value, na.rm = TRUE)) |
      (Goal == "max" & value == max(value, na.rm = TRUE))
    ) %>%
    slice_max(order_by = Confidence, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    dplyr::select(Metric, Goal, Confidence, value, Weight) %>%
    dplyr::mutate(`Weighted Confidence` = Confidence * Weight)
  optimal_confidence = sum(best_confidence_values$`Weighted Confidence`)

  optimal_values <- metrics_group %>%
    dplyr::filter(Confidence == round(optimal_confidence, digits = 1))
  precision_value <- round(as.numeric(optimal_values$Precision[1]), digits = 3)
  accuracy_value <- round(as.numeric(optimal_values$Accuracy[1]), digits = 3)
}
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, results='asis', fig.height = 6, fig.width = 8}
# printing out a table of the difference metrics with their respective goals and optimal confidence thresholds
if (nrow(sub_false) <= 10 & sum(pa_group$`Percent Agreement`, na.rm = T) == 0) {
  cat("\n\n**No performance metrics or difference distribution data available for this species.**\n\n")
  did_we_skip <- "yes"
} else {
cat(paste0("\n\nThe table below represents the optimal confidence thresholds for each critical parameter. An optimal confidence threshold was selected for each parameter based on the relative 'goals' for that parameter (e.g., selecting the threshold where ACV was at a minimum). Each parameter was then given a weight that was multiplied by each selected confidence. The sum of the weighted confidence variable indicates that across all years ", round(as.numeric(optimal_confidence), digits = 2), " is the optimal threshold for ", spec_pretty, ". The precision and accuracy values that correspond to the rounded optimal confidence threshold are ", precision_value, " and ", accuracy_value, ", respectively.\n"))
mt <- flextable(best_confidence_values) %>% autofit() %>% 
colformat_num(j = c("value"), digits = 4) %>% fit_to_width(8)
cat(knitr::knit_print(mt))

## building the line graph for all parameters
params_plot_dat <- params_plot_dat %>%
  dplyr::filter(Metric %in% c("Avg False Negatives", "Avg False Positives", "ACV", "PercAgree"))
colors <- c("Avg False Negatives" = "#228B22", "Avg False Positives" = "#af2634", "ACV" = "#FFD580", "PercAgree" = "#4E79A7")


# Building the difference distribution plot
params_plot <- ggplot(params_plot_dat, aes(x = as.numeric(Confidence), y = value, color = Metric)) +
  geom_rect(aes(xmin = optimal_confidence - 0.02, xmax = optimal_confidence + 0.02, ymin = -Inf, ymax = Inf, show.legend = F), color = "white", fill = "#FF474C", alpha = 0.01, show.legend = F, inherit.aes = FALSE) +
  geom_line(linewidth = 1.75) + 
  scale_color_manual(values = colors) +
  labs(
    title = "VIAME Performance Metrics Across all Years",
    caption = "Shaded red bar represents the optimal confidence threshold\nas determined by the weighting values illustrated in the table above.",
    x = "Confidence",
    y = "Value",
    color = "Metric"
  ) +
  scale_x_continuous(breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95), expand = c(0,0)) +
  theme_bw() +
  theme(plot.title = element_text(vjust = 2, size = 14, hjust = 0.5, face = "bold", family = "Times", color = "#5A5F63"),
        plot.subtitle = element_text(vjust = 2, size = 12, hjust = 0.5, family = "Times", face = "italic", color = "#5A5F63"),
        axis.text.x=element_text(size=12, family = "Times", vjust = 0.5, angle = 45),
        axis.text.y=element_text(size=12, family = "Times"),
        axis.title.x=element_text(size=14, face="bold", family = "Times", vjust = -3, color = "#5A5F63"),
        axis.title.y=element_text(size=14, face="bold", family = "Times", vjust = 4, lineheight = 0.5, color = "#5A5F63"),
        legend.title = element_text(size = 12, family = "Times", color = "#5A5F63", face = "bold"),
        legend.text = element_text(size = 10, family = "Times", color = "#5A5F63"),
        legend.position = "right",
        plot.caption = element_text(size = 10, family = "Times", face = "italic", hjust = 0, vjust = -5),
        panel.grid.major.y = element_line(color="#D9DADB", linewidth = .2),
        panel.grid.minor.y = element_blank(),
        #panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.spacing=unit(0.3,"lines"),
        strip.background=element_rect(fill = "grey60", color = "grey60"),
        axis.line = element_line( color="#5A5F63", linewidth = .2),
        axis.text = element_text(family = "Times", 
                                 colour = "#5A5F63", 
                                 size = 11),
        panel.background = element_blank(),
        plot.margin = unit(c(0.2,0.2,0.3,0.3), "inches")) +
  guides(
    color = guide_legend(position = "right", override.aes = list(size = 5), theme = theme(legend.key.spacing.x = unit(0.5, "cm")))
  )
print(params_plot)
ggsave(params_plot, filename = paste0(outdir, "Part III - Data Analysis/Figures/", species, "_performance_metrics.png"), device = "png", dpi = 300, height = 6, width = 8)
}
```

***
    
# Manual and VIAME Count Distribution
Deployment-level model and manual `REFERENCE` MaxN counts are compared by analyzing the "spread" across all confidence thresholds. Dot size corresponds to frequency of occurrence of a difference value, while color indicates whether the model MaxN was below, agreed with, or above the reference MaxN. The shaded red bar represents the confidence threshold at which the standard deviation was the lowest for **weighted differences (difference * frequency)**. If no shaded bar is present then the frequency was too low for a standard deviation to be calculated for all confidence thresholds or there was no single-best confidence threshold (i.e., the lowest standard deviation was the same for multiple thresholds).
```{r, echo=FALSE, warning=FALSE, error=FALSE, results='asis', fig.height = 6, fig.width = 8}
if (did_we_skip != "yes"){
    diff_dist <- spec_master %>% 
      dplyr::mutate(Difference = VIAME_MaxN - Manual) %>%
      dplyr::group_by(Species, Confidence, year, Version, Model_info, Difference) %>% 
      dplyr::summarise(Freq = n(), .groups = 'drop') %>%
      dplyr::mutate(Difference_weighted = Difference * Freq)

    plot_data <- diff_dist %>%
      dplyr::mutate(Difference_Type = factor(case_when(
        Difference < 0 ~ "Under-count (AI < Manual)",
        Difference == 0 ~ "Correct (AI = Manual)",
        Difference > 0 ~ "Over-count (AI > Manual)"), levels = c("Under-count (AI < Manual)", "Correct (AI = Manual)", "Over-count (AI > Manual)")))
    
    my_colors <- c(
      "Under-count (AI < Manual)" = "#006d2c",  
      "Correct (AI = Manual)"     = "#c7e9c0", 
      "Over-count (AI > Manual)"  = "#74c476")
    
#### Calculating spread across all years through standard deviation and then selecting the confidence threshold with the lowest spread. This is done using the difference_weighted variable which takes into account the frequency of occurrence of optimal or sub-optimal values.
spread <- plot_data %>%
  dplyr::group_by(Confidence) %>%
  dplyr::summarise(SD = sd(Difference_weighted)) %>%
  arrange(SD, .desc = T) # arrange by lowest spread 
bar_position <- as.numeric(factor(plot_data$Confidence))[plot_data$Confidence == as.character(spread[[1,1]])][1]   
bar_position <- ifelse(is.na(spread[[1,2]]) | (spread[[1,2]] == spread[[2,2]]), NA, bar_position)

difference_plot <- ggplot(plot_data, aes(x = as.character(Confidence), y = Difference, color = Difference_Type)) +
  geom_rect(aes(xmin = bar_position - 0.5, xmax = bar_position + 0.5, ymin = -Inf, ymax = Inf, show.legend = F), color = "white", fill = "#FF474C", alpha = 0.01, show.legend = F, inherit.aes = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  geom_point(aes(size = Freq, x = as.character(Confidence), y = Difference)) + 
  facet_wrap(~year, scales = "free_y", nrow = 1) +
  scale_color_manual(values = my_colors, name = "") +
  labs(
    title = "Distribution of Count Differences (AI vs. Manual)",
    caption = "",
    x = "Confidence",
    y = "Difference (AI - Manual)",
    fill = "Comparison"
  ) +
  theme_bw() +
  theme(plot.title = element_text(vjust = 2, size = 14, hjust = 0.5, face = "bold", family = "Times", color = "#5A5F63"),
        plot.subtitle = element_text(vjust = 2, size = 12, hjust = 0.5, family = "Times", face = "italic", color = "#5A5F63"),
        strip.text.y = element_text(size = 12, family = "Times", color = "white", angle = 0),
        strip.text.x = element_text(size = 12, family = "Times", color = "white", angle = 0),
        axis.text.x=element_text(size=12, family = "Times", vjust = 0.5, angle = 90),
        axis.text.y=element_text(size=12, family = "Times"),
        axis.title.x=element_text(size=14, face="bold", family = "Times", vjust = -3, color = "#5A5F63"),
        axis.title.y=element_text(size=14, face="bold", family = "Times", vjust = 4, lineheight = 0.5, color = "#5A5F63"),
        legend.title = element_text(size = 14, family = "Times", color = "#5A5F63"),
        legend.text = element_text(size = 12, family = "Times", color = "#5A5F63"),
        legend.position = "right",
        plot.caption = element_text(size = 12, family = "Times", face = "italic", hjust = 0, vjust = -3),
        panel.grid.major.y = element_line(color="#D9DADB", linewidth = .2),
        panel.grid.minor.y = element_blank(),
        #panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.spacing=unit(0.3,"lines"),
        strip.background=element_rect(fill = "grey60", color = "grey60"),
        axis.line = element_line( color="#5A5F63", linewidth = .2),
        axis.text = element_text(family = "Times", 
                                 colour = "#5A5F63", 
                                 size = 11),
        panel.background = element_blank(),
        plot.margin = unit(c(0.2,0.2,0.3,0.3), "inches")) +
  guides(
    color = guide_legend(position = "bottom", override.aes = list(size = 5), theme = theme(legend.key.spacing.x = unit(0.5, "cm")))
  )
print(difference_plot)
ggsave(difference_plot, filename = paste0(outdir, "Part III - Data Analysis/Figures/", species, "_count_differences.png"), device = "png", dpi = 300, height = 6, width = 8)
}
```


***

<br> 


```{r, echo=FALSE, warning=FALSE, error=FALSE, results = 'asis', message = FALSE, fig.height = 5.5, fig.width = 7}
if (nrow(spec_master) == 0) {
    cat("\n\n**No further data will be generated for this species due to no data available. A lack of data could be due to the filtering out of large schools (i.e., schools manually coded as 299, 399, or 999).**\n\n")
  precision <- list()
  difference <- list()
  false_ratios <- list()
  precision[[1]] <- data.frame(n = 0, validn = 0, R = NA, PercAgree = NA, ASD = NA, ACV = NA, AAD = NA, APE = NA, Confidence = "NONE", year = "NONE")
  difference[[1]] <- data.frame(Count = NA, Frequency = NA, Confidence = "NONE", year = "NONE")
  } else {
cat("\n# Analysis by Year {.tabset}\n\n")
# Creating the blank datasets that will get outputted from the loop. These datasets will contain both confidence by year information for each species
precision <- list()
difference <- list()
false_ratios <- list()

# running the big year over year loop
for (year.run in unique_years) {
# subsetting the data    
pa_spec_year <- pa_spec %>% dplyr::filter(year == year.run)
spec_master_year <- spec_master %>% dplyr::filter(year == year.run)
sub_false_year <- sub_false %>% dplyr::filter(year == year.run)
mod_false <- sub_false_year %>% 
  dplyr::mutate(Species = gsub("_", " ", str_to_sentence(Species)),
                year = as.character(year)) %>%
  dplyr::select(year, Confidence, Precision, Accuracy, FP, FN, False_P_Ratio, False_N_Ratio, FPR, FNR, Total_Actual_Positives, Total_Actual_Negatives)

mod_false_table <- mod_false %>%
  dplyr::group_by(year, Confidence) %>%
  summarise_all(mean)


# Create the top-level Year Tab 
cat(paste0("\n\n## ", year.run, "\n"))


#### FALSE POSITIVES AND NEGATIVES
cat("\n### False Positives and Negatives\n")

if (nrow(mod_false) == 0) {
    cat("No false positives or negatives for this species.\n\n")
  } else {
    cat("Average VIAME performance metrics for each confidence threshold.\n\n")
    ft <- flextable(mod_false_table) %>% autofit() %>% 
          colformat_num(j = c("False_P_Ratio", "False_N_Ratio"), digits = 4) %>% 
          fit_to_width(9)
    cat(knitr::knit_print(ft))
  }
 
cat("\n\n### Linear Model Plots for all Confidences\n")

if (nrow(spec_master_year) > 0) {
# Call the function to save and print the plot
print(lm_plots(spec_master_year, species, year.run))
} else {
  cat("\n\n**No linear model plots for this species for this year due to a lack of data.**\n\n")
}

    #### exact vs. relaxed agreement
    if (nrow(pa_spec_year) > 0) { # only need this tab if there is data for percent agreement
    cat("\n\n### ", "Exact vs. Relaxed Agreement", "\n") 
    cat("**Exact percent agreement** is the percent of deployments where the manual `REFERENCE` MaxN and model MaxN matched exactly.\n\n")
    cat("**Relaxed percent agreement** is the percent of deployments where the manual `REFERENCE` MaxN and model MaxN were within +/- 1 of each other.\n\n")
  PA_comp_plot <- ggplot(pa_spec_year, aes(x=as.factor(Confidence), y=percentage_1s, fill=Agreement)) +
  geom_bar(stat="identity", position="dodge", alpha=0.7, color="black") +
  labs(title=" ",
       x="Confidence",
       y="% Deployments") +
  scale_fill_manual(values=c("navyblue", "aquamarine")) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  theme_bw() +
  theme(plot.title = element_text(vjust = 2, size = 14, hjust = 0.5, face = "bold", family = "Times", color = "#5A5F63"),
        plot.subtitle = element_text(vjust = 2, size = 12, hjust = 0.5, family = "Times", face = "italic", color = "#5A5F63"),
        strip.text.y = element_text(size = 12, family = "Times", color = "white", angle = 0),
        strip.text.x = element_text(size = 12, family = "Times", color = "white", angle = 0),
        axis.text.x=element_text(size=12, family = "Times", vjust = 0.5, angle = 90),
        axis.text.y=element_text(size=12, family = "Times"),
        axis.title.x=element_text(size=14, face="bold", family = "Times", vjust = -3, color = "#5A5F63"),
        legend.key.spacing.x = unit(1.0, "cm"),
        axis.title.y=element_text(size=14, face="bold", family = "Times", vjust = 4, lineheight = 0.5, color = "#5A5F63"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12, family = "Times", color = "#5A5F63"),
        legend.position = "top",
        plot.caption = element_text(size = 12, family = "Times", face = "italic", hjust = 0, vjust = -3),
        panel.grid.major.y = element_line(color="#D9DADB", linewidth = .2),
        panel.grid.minor.y = element_blank(),
        #panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.spacing=unit(0.3,"lines"),
        strip.background=element_rect(fill = "grey60", color = "grey60"),
        axis.line = element_line( color="#5A5F63", linewidth = .2),
        axis.text = element_text(family = "Times", 
                                 colour = "#5A5F63", 
                                 size = 11),
        panel.background = element_blank(),
        plot.margin = unit(c(0.2,0.2,0.3,0.3), "inches"))
    print(PA_comp_plot)
    }


cat("<div class='panel panel-default' markdown='1'>")
cat("<div class='panel-heading'>")
cat("</div>")

cat("<div class='panel-body' markdown='1'>")
  
# Create the 'Confidence' sub-tab 
cat("\n\n### Confidence Thresholds {.tabset .tabset-fade .tabset-pills}\n")
  
# Confidence loop 
for (c in c(1:9,95)) {
  conf = as.numeric(paste0("0.", c))
  loop_key <- paste(year.run, conf, sep = "_")
  
  cat(paste0("#### ", conf, "\n")) # Creates the 0.1, 0.2 tabs
  
  subdf <- spec_master_year %>% dplyr::filter(Confidence == conf)
  
  if (nrow(subdf) == 0) {
    cat(paste0("No presence at a confidence of ", conf, " for ", gsub("_", " ", str_to_sentence(species)), ".\n\n"))
    
    # creating dummy precision and difference data frames to be joined with the rest when there is actual data available
    precision[[loop_key]] <- data.frame(n = 0, validn = 0, R = NA, PercAgree = NA, ASD = NA, ACV = NA, AAD = NA, APE = NA, Confidence = conf, year = year.run)
    difference[[loop_key]] <- data.frame(Count = NA, Frequency = NA, Confidence = conf, year = year.run)
    
  } else if (nrow(subdf) == 1) {
    
    #### bias
    cat("\n\n##### ", "Bias", "\n")    
    ab.out = ageBias(VIAME_MaxN~Manual, data = subdf, nref.lab="VIAME MaxN", ref.lab="Validated Set MaxN")
    summary(ab.out,what=c("n"))
    sink(tempfile())
    sym1 <- summary(ab.out,what=c("table",flip.table=FALSE, na.rm = T))
    sink()
    if (nrow(sym1)>=40){
    sym2 <- data.frame(array(sym1, dim(sym1)))
    names(sym2) <- as.numeric(row.names(sym2))-1
    row.names(sym2) <- as.numeric(row.names(sym2))-1
    sym2 <- sym2[colSums(sym2) != 0]
    sym2 <- sym2[rowSums(sym2) != 0,]
    cat(pander(sym2, style = "rmarkdown"))
    } else {
    cat(pander(sym1, style = "rmarkdown"))      
    }
    sink(tempfile())
    sym3 <- data.frame(summary(ab.out,what=c("symmetry")))
    sink()
    cat(pander(sym3, style = "rmarkdown"))
    plot(ab.out)
    plotAB(ab.out)
    
    cat("\n\n##### ", "Precision, Difference, and Models", "\n") 
    cat(paste0("Only one data point for a confidence of ", conf, ". Therefore, no precision, difference, or model output summaries.\n\n"))
    
    # creating dummy precision and difference data frames to be joined with the rest when there is actual data available
    precision[[loop_key]] <- data.frame(n = 1, validn = 1, R = NA, PercAgree = NA, ASD = NA, ACV = NA, AAD = NA, APE = NA, Confidence = conf, year = year.run)
    difference[[loop_key]] <- data.frame(Count = NA, Frequency = NA, Confidence = conf, year = year.run)
  } else {

    
    #### bias
    cat("\n\n##### ", "Bias", "\n")    
    ab.out = ageBias(VIAME_MaxN~Manual, data = subdf, nref.lab="VIAME MaxN", ref.lab="Validated Set MaxN")
    
    cat("\n\n**Figure 1**. Shows potential differences in paired counts between validated counts and VIAME counts. The mean (points) and 95% confidence interval (vertical lines) for the VIAME counts are computed at each value of the validated counts. The dotted line represents a 1:1 agreement line between the two counts.\n")
    plotAB(ab.out)
    cat("\n\n")  
    
    cat("\n\n**Figure 2**. Plots the difference between the two counts on the y-axis against the validated counts on the x-axis. A histogram of the validated counts is shown in the top margin and a histogram of the differences if shown in the right margin. The horizontal line represents an agreement line at a difference of zero. Also shown are the means and ranges for the VIAME counts at each of the validated counts (points and vertical lines, respectively). Means with an open dot are mean count differences that are significantly different from zero.\n")
    plot(ab.out)
    cat("\n\n")   
    
    summary(ab.out,what=c("n"))
    sink(tempfile())
    sym1 <- summary(ab.out,what=c("table",flip.table=FALSE))
    sink()
    sym2 <- data.frame(array(sym1, dim(sym1)))
    names(sym2) <- as.numeric(row.names(sym2))-1
    row.names(sym2) <- as.numeric(row.names(sym2))-1
    sym2 <- sym2[colSums(sym2) != 0]
    sym2 <- sym2[rowSums(sym2) != 0,]
    cat('\n')
    cat(pander(sym2, style = "rmarkdown"))
    sink(tempfile())
    sym3 <- data.frame(summary(ab.out,what=c("symmetry")))
    sink()
    
    cat('\n')
    cat(pander(sym3, style = "rmarkdown"))
    cat("\n\n")


cat('\n')
cat("\n\n")

    #### precision
    cat("\n\n##### ", "Precision", "\n") 
    ap.out <- agePrecision(~Manual+VIAME_MaxN,data=subdf)
    sink(tempfile())
    precdetails <- summary(ap.out,what="details") 
    precdetails <- precdetails[order(precdetails$CV),]
    precdetails <- precdetails[order(precdetails$PE, decreasing = T),]
    sink()
    #cat(pander(precdetails, style = "rmarkdown"))
    sink(tempfile())
    precision[[loop_key]] <- data.frame(summary(ap.out,what="precision"))
    sink()
    cat(pander(precision[[loop_key]], style = "rmarkdown"))
    precision[[loop_key]] <- precision[[loop_key]] %>% dplyr::mutate(Confidence = conf, year = year.run)
    
    cat("\n\n")
    cat("\n\n")
    
    
    #### difference
    cat("\n\n##### ", "Difference", "\n") 
    sink(tempfile())
    difference[[loop_key]] <- data.frame(summary(ap.out,what="difference"))
    sink()
    if (length(difference[[loop_key]]) > 2) {
      difference[[loop_key]] <- difference[[loop_key]] %>% dplyr::select(Var2, Freq)
      names(difference[[loop_key]]) <- c("Count", "Frequency")
    } else {
      names(difference[[loop_key]]) <- c("Count", "Frequency")
      difference[[loop_key]] <- difference[[loop_key]] %>% dplyr::filter(Frequency > 0)
    }
    cat(pander(difference[[loop_key]], style = "rmarkdown"))
    difference[[loop_key]] <- difference[[loop_key]] %>% dplyr::mutate(Confidence = conf, year = year.run)
    cat("\n\n&nbsp;\n")

    cat("\n\n")
    cat("\n\n")
    
    #### models
    cat("\n\n##### ", "Models", "\n") 
    glm.out <- glm(Manual ~ VIAME_MaxN, family = poisson(),data=subdf)
    sink(tempfile())
    glmsummary <- summary(glm.out)
    sink()
    print(glmsummary$call)
    cat(pander(glmsummary))
    cat("\n\n&nbsp;\n")
    cat("\n\n&nbsp;\n")
    lm.out <- lm(Manual ~ VIAME_MaxN,data=subdf)
    sink(tempfile())
    lmsummary <- summary(lm.out)
    sink()
    print(lmsummary$call)
    cat(pander(lmsummary))
    cat("\n\n")
    cat("\n\n")
  }
 }
  cat("</div>")
  cat("</div>")
}}
```
